---
title: "Data dictionary"
format: html
editor: visual
---

Access function: [Data dictionary](Scripts/Data dictionary.R). 

## Set up and load data

```{r packages, collapse = T}

library(haven)
library(tidyverse)
library(labelled)
library(huxtable)
library(openxlsx)

```

```{r load data}

# Load BSA teaching data
df <- read_sav('/Users/joecrowley/R/Data/BSA Teaching Data/spss/spss25/bsa2019_poverty_open.sav', user_na = T)

df %>% 
  select(welfare2, leftrigh, libauth, RSex, RAgeCat) %>% 
  look_for()

```

<br> 

## unique_vals

```{r}

unique_vals <- function(data, max_values = 30){
  
  vect <- 
    data %>% 
    map(., ~unique(to_factor(.x))) %>% 
    map(., function(xxx) {
      
      vec_length <- length(xxx)
      if(vec_length > max_values){vec_length <- max_values}
      result <- paste0(xxx[1:vec_length], collapse = ", ")
      return(result)
      
      }) %>%
     as_tibble %>% 
    pivot_longer(cols= everything())
  
  count_unique <- data %>% 
    map(., ~length(unique(to_factor(.x)))) %>% 
    as_tibble() %>% 
    pivot_longer(cols = everything()) %>% 
    rename(n_unique = value)
  
  vect <- full_join(count_unique, vect, by = "name") 
  names(vect)[3] <- paste("First", max_values, "unique values")
  
  return(vect)

}

```

<br> 

### Trials

#### Basic

```{r}

unique_vals(df)

```

<br> 

#### Basic - with flextable

```{r}

unique_vals(df) %>% flextable::flextable() %>% flextable::autofit()

```

<br> 

#### Set number of unique values to show

```{r}

unique_vals(df, max_values = 10) %>% flextable::flextable() %>% flextable::autofit()

```

<br> 

## df_count 

```{r count data frame}

df_count <- function(data) { 
    map(names(data), ~data %>% 
          pull(all_of(.x)) %>% 
          to_factor() %>% 
          fct_count() %>% 
          mutate(name = .x, .before = 1) %>% 
          rename(levels = f) %>% 
          filter(n != 0) %>%
          mutate(id = row_number()) %>%
          filter(id %in% 1:10) %>% 
          mutate(Count = paste0(levels, ";  N = ", n)) %>% 
          select(-id, -n, -levels)
        ) %>% 
      bind_rows 
}

```

<br> 

### Trials


```{r}

df %>% df_count

```

<br> 

```{r}

df %>% df_count() %>% print(n=100) %>% flextable::flextable() %>% flextable::autofit()

```

<br> 

## fullsome

```{r}

fullsome <- function(data, max_values = 30, count = F, unique_vals = T){
  
  components <- list(
  
    map(data, ~sum(is.na(.x))) %>% as_tibble() %>% pivot_longer(cols = everything()) %>% rename(n_miss = value) %>% mutate(p_miss = paste0(round(n_miss / nrow(data) * 100, 2), "%")),
    
    map(data, ~paste0(class(.x), collapse = ", ")) %>% as_tibble() %>% pivot_longer(cols = everything()) %>% rename(class = value),
    
    map(data, ~var_label(.x)) %>% as_tibble() %>% pivot_longer(cols = everything()) %>% rename(label = value)
    
    )
  
  if(unique_vals == T){
    components <- append(components, list(unique_vals(data, max_values = max_values)))
  }
  
  results <- reduce(components, full_join, by = "name") %>% 
    relocate(name, label, any_of("n_unique"), n_miss, p_miss, class)
  
    if (count == T) {
    
    counts <- data %>% df_count()
    results <- full_join(results, counts, by = "name")
        
  }
  
  return(results)
  
}

```

<br> 

### Trials

<br> 

#### Basic

```{r}

fullsome(df)

```

<br> 

#### With flextable + number of values to show

```{r}

fullsome(df, max_values = 10) %>% flextable::flextable() %>% flextable::autofit()

```

<br> 

#### Simplified

No unique values, no count. Also - non-disclosive. 

```{r}

simplified <- fullsome(df, max_values = 10, unique_vals = F) 
simplified%>% flextable::flextable() %>% flextable::autofit()
```

<br> 

#### With counts

```{r}

counter <- fullsome(df, max_values = 10, count = T) 

```

<br> 

#### With counts

```{r}

counter2 <- fullsome(df, max_values = 10, count = T, unique_vals = F) 
counter2 %>% flextable::flextable() %>% flextable::theme_zebra()

```

<br> 

## Bespoke huxtable formatting for fullsome()

```{r}

overview_hux <- function(data, flex = F){
  
   number_of_cols <- ncol(data)
  
    # Convert to huxtable
    ht <- as_hux(data)
    
    # Vertically center-align all cells
    ht <- set_valign(ht, everywhere, 1:number_of_cols, "middle")
    
    # Assign a fill color per group
    group_ids <- data$name # Unique IDs for groups are the variable names
    change <- group_ids == lag(group_ids) # Does variable name change? 
    change[1] <- TRUE # Replace a missing value
    change
    group_colours <- factor(c("lightblue", "lightgreen")) # Row colour palette
    group_colours
    # For length of table identify the colour to use in each row
    colours <- vector()
    for(i in 1:length(group_ids)) {
      # Each time the variable name changes reverse the colour palette
      if(change[i] == FALSE){group_colours <- rev(group_colours)}
      # Build up a vector of colours to use. Convert to character otherwise becomes numeric.
      colours <- append(colours, as.character(group_colours[1]))
    }
    colours
    
    # Set the colours to be used
    ht <- set_background_color(ht, row = 2:nrow(ht), col = 1:ncol(ht), colours)
    ht <- set_background_color(ht, row = 1, col = everywhere, "lightgrey")
    
    # Merge repeated rows in the first seven columns
    # Only do so where counts are added
    if("Count" %in% names(data)){
      r <- rle(group_ids) # Count lengths of runs of repeated values
      r <- r$lengths # Extract those lengths from resulting object
      merge_start <- 2 # Start on row 2
      
      # Loop over sets of repeated rows merging columns 1 to 7
      for(i in 1:length(r)) {
        ht <- set_rowspan(ht, row = merge_start, col = 1:(number_of_cols-1), value = r[i])
        merge_start <- merge_start + r[i]
        }
    }
    
    # Add borders and alignment
    ht <- ht |>
      set_all_borders(1) |>
      set_align(everywhere, everywhere, "center") |>
      set_width(0.8)
    
    # if flex is TRUE then return flextable
    if(flex == T){
    return(ht %>% as_flextable())
    } else {
      # Otherwise return the huxtable
      return(ht)
    }
  
}

```

### Trial

```{r}

counter %>% overview_hux(flex = T)

```

### Trial 2

```{r}

counter2 %>% overview_hux(flex = T)

```

### Trial 3

```{r}

simplified %>% overview_hux(flex = T)

```

## Export

```{r}

wb <- as_Workbook(counter %>% overview_hux(flex = F), sheet = "overview")
freezePane(wb, "overview", firstActiveRow = 2, firstActiveCol = NULL)
saveWorkbook(wb, file = "./Data dictionary example.xlsx", overwrite = T)

```

## Save out function 

```{r}
# Collect both functions as named list
fns <- list(
  df_count = df_count,
  unique_vals= unique_vals,
  fullsome = fullsome, 
  overview_hux = overview_hux
)

# Write each to the same file, separated by a blank line
writeLines(
  unlist(lapply(names(fns), function(nm) {
    c(paste0(nm, " <- "), deparse(fns[[nm]]), "")
  })),
  "/Users/joecrowley/R/Exploratory/Scripts/Data dictionary.R"
)
```








