---
title: "Export to PowerPoint"
format: html
editor: visual
---

## Overview

The slide-building helpers now live in the **`deck`** package, kept in this
repository under `deck/`. This page demonstrates every builder and writes
`Example PowerPoint from functions.pptx`.

Install the package once from the repo root:

```r
devtools::install("deck")
# or, without installing:  devtools::load_all("deck")
```

The functions were originally developed inline in this document and dumped to
`Scripts/ppt_functions/`; that folder is now superseded by the package.

-   Local copy: [Example PowerPoint from functions.pptx](`r URLencode("Example PowerPoint from functions.pptx")`)
-   GitHub copy: [Example PowerPoint from functions.pptx](https://raw.githubusercontent.com/Joeacrowley/Exploratory/main/Example%20PowerPoint%20from%20functions.pptx)

<br>

```{r packages, collapse = T}
library(tidyverse)
library(mschart)
library(officer)
library(flextable)
library(deck)
```

<br>

## Open the template

`mint_pptx()` opens the bundled "Sage Mint Theme" template and returns an
`officer` `rpptx` object. Every `*_slide()` builder takes that object as its
first argument, returns it invisibly, and can be chained with `|>`.

```{r}
ppt <- mint_pptx()
```

<br>

## Available layouts

`mint_layouts()` is the replacement for the old `available_styles()` /
`available_style_names()` helpers. It filters `officer::layout_properties()`
to the layouts the builders target and shows the real placeholder labels --
useful when a `ph_with()` call fails.

```{r}
mint_layouts(ppt)

mint_layouts(ppt, placeholders = FALSE)
```

<br>

## Title

```{r}
ppt <- ppt |>
  add_title_slide(
    title = "Title text example",
    subtitle = c("Can you have multiple strings", "in the subtitle?")
  )
# A subtitle vector with multiple elements is split over multiple lines.
```

<br>

## Divider slide

```{r}
ppt <- ppt |>
  add_divider_slide(
    text = "Title for divider slide",
    footer = c("Can you have multiple strings in the footer?",
               "Yes, but they'll appear off screen"),
    header = "Header"
  )
# The footer placeholder on this layout sits low and short, so a long or
# multi-line footer overflows the slide.
```

<br>

## Title and full width content

```{r}
ppt <- ppt |>
  full_width_content_slide(
    title = "Title",
    text = c("Test text", "test", "living"),
    footer = "Footer"
  )
```

<br>

## Title and two column content

```{r}
ppt <- ppt |>
  double_text_column_slide(
    title = "Title",
    left_text = c("Left_text", "test", "living"),
    right_text = c("Right text", "test", "living"),
    footer = "Footer"
  )
```

<br>

## Title and full width table with description

The template's table area is 8.42 cm tall and 30.86 cm wide. `make_flex()`
styles a data frame, and `full_width_table_slide()` shrinks it toward that
box with `fit_to_height()` + `flextable::fit_to_width()` unless
`control_height = FALSE`.

`fit_to_height()` is only a heuristic (trim padding, then force uniform row
heights) -- it cannot rescue an arbitrarily large table, and when rows have
to be compressed below their natural height it warns that the table will
still overflow in PowerPoint. The 15-row example below is deliberately too
tall, so that warning is expected here; pass `warn = FALSE` to silence it.

```{r}
ft_mtcars <- mtcars |>
  rownames_to_column() |>
  slice(1:15) |>
  flextable::flextable() |>
  flextable::theme_zebra() |>
  flextable::autofit()

# Standalone preview of the fitting helpers
ft_mtcars |>
  fit_to_height(max_height = 8.42 / 2.54) |>
  flextable::fit_to_width(max_width = 30.86, unit = "cm")
```

```{r}
ppt <- ppt |>
  full_width_table_slide(
    title = "Full width table slide",
    text = c("Description text",
             "Here table is shrunk a bit to fit",
             "if table too big, still wont' fit"),
    table = ft_mtcars,
    footer = "Footer text"
  ) |>
  full_width_table_slide(
    title = "Full width table slide",
    text = "Description text. Here is an unshrunk table, too big",
    table = ft_mtcars,
    footer = "Footer text",
    control_height = FALSE
  )
```

<br>

## A flextable function

`make_flex()` converts a data frame to a Sage Mint styled flextable: zebra
striping, thin interior rules, black lines under the header and at the foot
of the body, and fixed-digit number formatting. Position any further
flextable edits after it. Pass `caption` a string to add an auto-numbered
Word caption.

```{r}
mtcars |> rownames_to_column() |> slice(1:15) |> make_flex()
```

<br>

## Title and full width chart with description

```{r}
chart_example <- ms_barchart(
  data = mtcars |> rownames_to_column() |> slice(1:8),
  x = "rowname",
  y = "mpg"
)

ppt <- ppt |>
  full_width_chart_slide(
    title = "Title for full width chart with description",
    text = c("The description text", "spread over lines"),
    chart = chart_example,
    source = "Source text",
    footer = "Footer for chart"
  )
```

<br>

## Text on one side and chart on the other

```{r}
ppt <- ppt |>
  text_and_chart_slide(
    title = "Slide with one column of text and one chart",
    text = "The column of text",
    chart = chart_example,
    source = "Chart source",
    footer = "Footer for slide"
  )
```

<br>

## Save output here

```{r}
print(ppt, target = "Example PowerPoint from functions.pptx")
```
