---
title: "Export to PowerPoint"
format: html
editor: visual
---
  
  
```{r packages, collapse = T}
library(tidyverse)
library(mschart)
library(officer)
library(flextable)

functions_folder <- "/Users/joecrowley/R/Exploratory/Scripts/ppt_functions"
```

<br> 

## Creat empty PowerPoint

```{r}
create_empty_ppt <- function(file){
    read_pptx(file)
}

ppt <- create_empty_ppt(file = "/Users/joecrowley/R/Exploratory/my_template.pptx")

writeLines(
  c("create_empty_ppt <- ", # Assignment needs to be added manually
    deparse(create_empty_ppt)), # Extract function code
  paste0(functions_folder,"/create_empty_ppt.R")) # File name is just the function name

```

<br> 

## Names of available styles

```{r}
available_style_names <- function(slides){

  layout_properties(slides) %>%
    filter(grepl("mint", name, ignore.case = T) |
             name %in% c("Title and full width content",
                         "Title and two column content",
                         "Title and full width table with description",
                         "Title and full width chart with description",
                         "Text and chart with source"
             )) %>%
    select(name) %>% unique() %>% unlist %>% unname()

}

available_style_names(ppt)

writeLines(
  c("available_style_names <- ", # Assignment needs to be added manually
    deparse(available_style_names)), # Extract function code
  paste0(functions_folder,"/available_style_names.R")) # File name is just the function name

```

<br> 

## Available styles

This shows the name of the objects (text boxes etc) inside the slide templates.

```{r}
available_styles <- function(slides){

    layout_properties(slides) %>%
    filter(grepl("mint", name, ignore.case = T) |
             name %in% c(
               "Title: Green", 
               "Divider: Green", 
               "End message: Green",
               "Title and full width content",
               "Title and two column content",
               "Title and full width table with description",
               "Title and full width chart with description",
               "Text and chart with source"
             )) %>%
    select(name, type, type_idx, id, ph_label)

}

styles_to_use <- available_styles(ppt)
styles_to_use %>% filter(name == "Title: Green")


writeLines(
  c("available_styles <- ", # Assignment needs to be added manually
    deparse(available_styles)), # Extract function code
  paste0(functions_folder,"/available_styles.R")) # File name is just the function name

```

<br> 

## Title

```{r}
# title
styles_to_use %>% filter(name == "Title: Green")

add_title_slide <- function(slides, title, subtitle = NULL){

  slides <- slides |>
    add_slide(layout = "Title: Green", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Text Placeholder 9"))

  if(!is.null(subtitle)){  slides <- slides |>
    ph_with(subtitle, ph_location_label(ph_label = "Subtitle 2"))
  }

}

# Example...
ppt <- add_title_slide(
  ppt,
  title = "Title text example",
  subtitle = c("Can you have multiple strings","in the subtitle?"))
# You can, if your subtitle vector has multiple elements the resulting
# subtitle is split over multiple lines.


writeLines(
  c("add_title_slide <- ", # Assignment needs to be added manually
    deparse(add_title_slide)), # Extract function code
  paste0(functions_folder,"/add_title_slide.R")) # File name is just the function name
```

<br> 

## Divider slide

```{r}

styles_to_use %>% filter(name == "Divider: Green")

add_divider_slide <- function(slides, text, footer = NULL, header = NULL){

  slides <- slides |>
    add_slide(layout = "Divider: Green", master = "Sage Mint Theme") |>
    ph_with(text, ph_location_label(ph_label = "Text Placeholder 9"))

  if(!is.null(footer)){  slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Straight Connector 3"))
  }

  if(!is.null(header)){  slides <- slides |>
    ph_with(header, ph_location_label(ph_label = "Subtitle 2")) # alleged subtitle, not really.
  }

}

# Example...
ppt <- add_divider_slide(
  ppt,
  text = "Title for divider slide",
  footer = c("Can you have multiple strings in the footer?", "Yes, but they'll appear off screen"),
  header = "Header")

writeLines(
  c("add_divider_slide <- ", # Assignment needs to be added manually
    deparse(add_divider_slide)), # Extract function code
  paste0(functions_folder,"/add_divider_slide.R")) # File name is just the function name
```

<br> 

## Title and full width content


```{r}
styles_to_use %>% filter(name == "Title and full width content")

full_width_content_slide <- function(slides, title, text, footer = NULL){

  slides <- slides |>
    add_slide(layout = "Title and full width content", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Title 1")) |>
    ph_with(text, ph_location_label(ph_label = "Content Placeholder 2"))


  if(!is.null(footer)){  slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Footer"))
  }


}

# Weirdly, you don't seem to need to assign the output to an object for it to
# updated. Which is good. But how / why / what the fuck?
full_width_content_slide(
  slides = ppt,
  title = "Title",
  text = c("Test text", "test", "living"),
  footer = "Footer"
)


writeLines(
  c("full_width_content_slide <- ", # Assignment needs to be added manually
    deparse(full_width_content_slide)), # Extract function code
  paste0(functions_folder,"/full_width_content_slide.R")) # File name is just the function name

```

<br>

## Title and two column content

```{r}
styles_to_use %>% filter(name == "Title and two column content")

double_text_column_slide <- function(slides, title, left_text = NULL, right_text = NULL, footer = NULL){

  slides <- slides |>
    add_slide(layout = "Title and two column content", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Title 1"))

  if(!is.null(left_text)){  slides <- slides |>
    ph_with(left_text, ph_location_type(type = "body", type_idx = 3))
  }

  if(!is.null(right_text)){  slides <- slides |>
    ph_with(right_text, ph_location_type(type = "body", type_idx = 4))
  }

  if(!is.null(footer)){  slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Footer"))
  }

}

double_text_column_slide(
  slides = ppt,
  title = "Title",
  left_text = c("Left_text", "test", "living"),
  right_text = c("Right text", "test", "living"),
  footer = "Footer"
)

writeLines(
  c("double_text_column_slide <- ", # Assignment needs to be added manually
    deparse(double_text_column_slide)), # Extract function code
  paste0(functions_folder,"/double_text_column_slide.R")) # File name is just the function name


```

<br>

## Title and full width table with description

<br> 

### Supporting function to find table size  

```{r}

styles_to_use %>% filter(name == "Title and full width table with description")
# Max height of table space - 8.42cm, converted to 3.314961 inches in code below
# Max width of table space - 30.86cm

# use for big tables only really.
ft_mtcars <- mtcars %>% rownames_to_column() %>% slice(1:15) %>% flextable::flextable() %>% flextable::theme_zebra() %>% flextable::autofit()

ft_mtcars %>%
  flextable::fit_to_width(max_width = 30.86, unit = "cm") # But not height constraining equivalent

# Helper function to limit export height.
# This is a chatgpt functions, it requires measurement in inches, unclear to me why.
fit_to_height <- function(ft, max_height, add_header = TRUE) {

  # --- Step 1: compute "natural" table height in inches
  # flextable dims are in inches
  dims <- flextable::flextable_dim(ft)
  natural_height <- dims$height

  # --- Step 2: reduce padding if needed
  if (natural_height > max_height) {
    ft <- flextable::padding(ft, padding.top = 1, padding.bottom = 1, part = "all")

    # recompute dimensions after padding shrink
    dims <- flextable::flextable_dim(ft)
    natural_height <- dims$height
  }

  # Count rows (optionally include header)
  if (natural_height > max_height) {

    n_rows <- flextable::nrow_part(ft, part = "body")
    if (add_header) {
      n_rows <- n_rows + flextable::nrow_part(ft, part = "header")
    }

    # Compute uniform row height
    row_height <- max_height / n_rows

    # Apply height to all parts
    ft <- flextable::height_all(ft, height = row_height)
  }

  return(ft)
}


ft_mtcars %>%
   fit_to_height(max_height = 3.314961) %>%
   flextable::fit_to_width(max_width = 30.86, unit = "cm")

```

<br> 

### Actual function to pass table to slide

```{r}
full_width_table_slide <- function(slides, title, text, table, footer = NULL, control_height = T){

  if(control_height == T){
  table_height_managed <- table %>%
    fit_to_height(max_height = 3.314961) %>%
    flextable::fit_to_width(max_width = 30.86, unit = "cm")
  } else {table_height_managed <- table}

  slides <- slides |>
    add_slide(layout = "Title and full width table with description", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Title 1")) |>
    ph_with(text, ph_location_type(type = "body", type_idx = 2)) |>
    ph_with(table_height_managed, ph_location_type(type = "tbl"))

  if(!is.null(footer)){ slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Footer"))
  }

}

full_width_table_slide(
  slides = ppt,
  title = "Full width table slide",
  text = c("Description text","Here table is shrunk a bit to fit", "if table too big, still wont' fit"),
  table = ft_mtcars,
  footer = "Footer text"
)

full_width_table_slide(
  slides = ppt,
  title = "Full width table slide",
  text = "Description text. Here is an unshrunk table, too big",
  table = ft_mtcars,
  footer = "Footer text",
  control_height = F
)

writeLines(
  c("full_width_table_slide <- ", # Assignment needs to be added manually
    deparse(full_width_table_slide)), # Extract function code
  paste0(functions_folder,"/full_width_table_slide.R")) # File name is just the function name

writeLines(
  c("fit_to_height <- ", # Assignment needs to be added manually
    deparse(fit_to_height)), # Extract function code
  paste0(functions_folder,"/fit_to_height.R")) # File name is just the function name

```

<br> 

## A flextable function

```{r}

# And also a flextable set up function...
# Converts a dataframe to a flextable
# Position other flextable edits after this...

# Set caption to a character string with the title.
# make_flex <- function(df, caption = F, digits = 0) {
#
#   table_width <- ncol(df)
#
#   ft <- df %>%
#     flextable::flextable() %>%
#     flextable::set_table_properties(layout = 'autofit', width = 1) %>%
#     flextable::theme_zebra(
#       odd_header = "grey90",
#       odd_body = "white",
#       even_body = "#f3f8fc"   # very gentle light blue
#     ) %>%
#       flextable::hline(i = 1, part = "header", border = fp_border(color = "black", width = 1)) %>%
#       flextable::hline_bottom(part = "body", border = fp_border(color = "black", width = 1)) %>%
#     flextable::colformat_double(digits = digits)
#
#   if(caption != F){
#     ft <- ft %>% flextable::add_header_row(values = caption, colwidths = table_width)
#   }
#
#   return(ft)
#
# }

make_flex <- function(df, caption = F, digits = 0) {

  table_width <- ncol(df)

  ft <- df %>%
    flextable::flextable() %>%
    flextable::set_table_properties(layout = 'autofit', width = 1) %>%
    flextable::theme_zebra(
      odd_header = "grey90",
      odd_body = "white",
      even_body = "#f3f8fc"   # very gentle light blue
    ) %>%
    border(part = "body", border = fp_border(color = "grey80", width = 1)) %>%
    flextable::hline(i = 1, part = "header", border = fp_border(color = "black", width = 1)) %>%
    flextable::hline_bottom(part = "body", border = fp_border(color = "black", width = 1)) %>%
    flextable::colformat_double(digits = digits)

  if(caption != F){
    ft <- ft %>% flextable::set_caption(caption = caption,
                                        autonum = run_autonum(seq_id = "tab"),
                                        word_stylename = "Table: Heading row_"
                                        )
  }

  return(ft)

}

mtcars %>% rownames_to_column() %>% slice(1:15) %>% make_flex()


writeLines(
  c("make_flex <- ", # Assignment needs to be added manually
    deparse(make_flex)), # Extract function code
  paste0(functions_folder,"/make_flex.R")) # File name is just the function name

```

<br> 

## Title and full width chart with description

```{r}

styles_to_use %>% filter(name == "Title and full width chart with description")

# Create a bar chart
chart_example <- ms_barchart(
  data = mtcars %>% rownames_to_column() %>% slice(1:8),
  x = "rowname",
  y = "mpg"
)

full_width_chart_slide <- function(slides, title, text = NULL, chart, source = NULL, footer = NULL){

  slides <- slides |>
    add_slide(layout = "Title and full width chart with description", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Title 1")) |>
    ph_with(chart, ph_location_type(type = "chart"))

  if(!is.null(text)){slides <- slides |>
    ph_with(text, ph_location_type(type = "body", type_idx = 2))
  }

  if(!is.null(source)){slides <- slides |>
    ph_with(source, ph_location_type(type = "body", type_idx = 3))
  }

  if(!is.null(footer)){slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Footer"))
  }

}

full_width_chart_slide(
  slides = ppt,
  title = "Title for full width chart with description",
  text = c("The description text", "spread over lines"),
  chart = chart_example,
  source = "Source text",
  footer = "Footer for chart"
)

writeLines(
  c("full_width_chart_slide <- ", # Assignment needs to be added manually
    deparse(full_width_chart_slide)), # Extract function code
  paste0(functions_folder,"/full_width_chart_slide.R")) # File name is just the function name

```

<br> 

## Text on one side and chart on the other

```{r}

styles_to_use %>% filter(name == "Text and chart with source")

text_and_chart_slide <- function(slides, title, text, chart, source = NULL, footer = NULL){

  slides <- slides |>
    add_slide(layout = "Text and chart with source", master = "Sage Mint Theme") |>
    ph_with(title, ph_location_label(ph_label = "Title 1")) |>
    ph_with(text, ph_location_type(type = "body", type_idx = 2), ) |>
    ph_with(chart, ph_location_type(type = "chart"))

  if(!is.null(source)){slides <- slides |>
    ph_with(source, ph_location_type(type = "body", type_idx = 3))
  }

  if(!is.null(footer)){slides <- slides |>
    ph_with(footer, ph_location_label(ph_label = "Footer"))
  }

}

text_and_chart_slide(
  slides = ppt,
  title = "Slide with one column of text and one chart",
  text = "The column of text",
  chart = chart_example,
  source = "Chart source",
  footer = "Footer for slide"
)

writeLines(
  c("text_and_chart_slide <- ", # Assignment needs to be added manually
    deparse(text_and_chart_slide)), # Extract function code
  paste0(functions_folder,"/text_and_chart_slide.R")) # File name is just the function name

```

<br> 

## Save output here

```{r}
print(ppt, target = "Example PowerPoint from functions.pptx")
```