---
title: "Frequency function for SPSS"
format: html
editor: visual
---

Access function: [Data dictionary](Scripts/freq_spss.R). 

## Package set up

```{r packages, collapse = T}
library(haven)
library(tidyverse)
library(labelled)
library(flextable)
library(huxtable)
```

```{r load data}

# Load BSA teaching data
df <- read_sav('/Users/joecrowley/R/Data/BSA Teaching Data/spss/spss25/bsa2019_poverty_open.sav', user_na = T)

df %>% 
  select(welfare2, leftrigh, libauth, RSex, RAgeCat) %>% 
  look_for()

```

<br> 

## freq_spss

### What the function does

`freq_spss()` produces a frequency table for a labelled survey variable, modelled on the output style of SPSS. It is designed for use with **haven-imported datasets**, where variables carry value labels, variable labels, and user-defined missing values (user NAs) — metadata that ordinary R frequency tools ignore.

The output distinguishes three kinds of rows:

- **Valid responses** — cases with a recognised, non-missing value. These rows show frequency, percentage of valid cases, and cumulative percentage within valid cases.
- **User-defined missing values** — codes the researcher has flagged as missing (e.g. "Don't know", "Refused") but which are tracked separately rather than silently dropped. These rows show frequency only.
- **Total columns** — percentage and cumulative percentage computed across *all* rows (valid + user NAs combined), replicating the "Total %" column familiar from SPSS output.

The result is either a plain **tibble** (for downstream manipulation) or a formatted **flextable** (for Word reports or RMarkdown output), depending on the `as_flextable` argument.

---

### How the code works

#### 1. Input validation and setup

The function begins with a simple data frame check, then extracts the target variable as a plain vector using `pull()`. A descriptive column label is constructed by combining the R variable name with the haven variable label (`var_label()`).

#### 2. Three factor views

The variable is converted into factors three times, each serving a different purpose:

| Object | Contents | How produced |
|---|---|---|
| `var_all` | All values including user NAs | `to_factor(levels = "prefixed")` |
| `var_no_NA` | Valid values only | `to_factor(..., user_na_to_na = TRUE)` |
| `var_only_NA` | User-NA values only | Subset raw variable to `na_values` attribute, then convert |

The `levels = "prefixed"` argument produces labels in the form `[1] Yes`, which preserves the numeric code alongside the label text.

#### 3. Three sub-tables, then joined

Rather than a single computation, the table is assembled from three separate tibbles that are then joined:

- **`valid_tbl`** — counts and proportions of non-missing cases, with cumulative % computed at this stage (so it runs only over valid rows).
- **`user_na_tbl`** — counts only, for user-NA rows.
- **`total_tbl`** — proportions across all rows combined.

These are joined sequentially with explicit `full_join(..., by = "f")` calls. The `n` and `type` columns that collide across joins are resolved with `coalesce()`. Rows are then `arrange()`d to put valid cases first before `cumsum()` computes the total cumulative percentage, ensuring stable ordering.

#### 4. Label deduplication

Haven's prefix format (`[1] Yes`) can produce redundant labels for unlabelled numeric variables, where a value like `1` gets labelled `[1] 1`. The function detects this by splitting each level on `] ` and comparing the left (value) and right (label) parts. Where they match, the prefix is stripped. A deduplication step guards against duplicate names in the recoding vector, which would otherwise trigger a warning from `fct_recode()`.

#### 5. Output

The `type` column and the generic `f` column name are tidied up — `f` is replaced by the descriptive variable label. If `as_flextable = TRUE`, the tibble is passed to an internal helper `.freq_format_table()`, which applies row shading (blue for valid, grey for user NAs), human-readable column headers, right-alignment, and vertical dividers.

*Requires: `dplyr`, `forcats`, `stringr`, `purrr`, `labelled`, `flextable`*

```{r freq_spss}

#' Generate an SPSS-style frequency table for a labelled variable
#'
#' @param data        A data frame containing the variable of interest.
#' @param var         The (unquoted) variable to tabulate.
#' @param as_flextable Logical. If TRUE, returns a formatted flextable;
#'                    otherwise returns a plain tibble. Default: FALSE.
#' @return A tibble or flextable with frequency, percentage, cumulative %,
#'         and totals, with user-defined missing values distinguished from
#'         valid values.

freq_spss <- function(data, var, as_flextable = FALSE) {
  
  # ── Input validation ─────────────────────────────────────────────────────────
  if (!is.data.frame(data)) stop("`data` must be a data frame.")
  
  # ── 1. Extract the raw variable and build a descriptive column label ──────────
  variable <- data %>% dplyr::pull({{var}})
  label <- paste0(
    data %>% dplyr::select({{var}}) %>% names,
    " - ",
    labelled::var_label(variable)
  )
  
  # ── 2. Build three views of the variable as factors ───────────────────────────
  var_all <- variable %>% labelled::to_factor(levels = "prefixed") %>% forcats::fct_drop()
  
  var_no_NA <- variable %>% labelled::to_factor(levels = "prefixed", user_na_to_na = TRUE) %>% forcats::fct_drop()
  
  user_na_vals <- attr(variable, "na_values")
  
  if (is.null(user_na_vals) || length(user_na_vals) == 0) {
    var_only_NA <- factor(character(0))
  } else {
    var_only_NA <- variable[variable %in% user_na_vals] %>%
      labelled::to_factor(levels = "prefixed") %>%
      forcats::fct_drop()
  }
  
  # ── 3. Build the frequency table by joining three sub-tables ──────────────────
  valid_tbl <- var_no_NA %>%
    stats::na.omit() %>%
    forcats::fct_count(prop = TRUE) %>%
    dplyr::filter(!is.na(f)) %>%
    dplyr::mutate(
      cum_p = paste0(round(cumsum(p) * 100, 2), "%"),
      p     = paste0(round(p * 100, 2), "%"),
      type  = "valid"
    )
  
  user_na_tbl <- var_only_NA %>%
    forcats::fct_count(prop = FALSE) %>%
    dplyr::mutate(type = "user_na")
  
  total_tbl <- var_all %>%
    forcats::fct_count(prop = TRUE) %>%
    dplyr::select(-n) %>%
    dplyr::rename(total_p = p)
  
  table <- valid_tbl %>%
    dplyr::full_join(user_na_tbl, by = "f") %>%
    dplyr::full_join(total_tbl,   by = "f") %>%
    dplyr::mutate(
      n    = dplyr::coalesce(n.x, n.y),
      type = dplyr::coalesce(type.x, type.y)
    ) %>%
    dplyr::select(-n.x, -n.y, -type.x, -type.y) %>%
    dplyr::arrange(desc(type == "valid"), f) %>%
    dplyr::mutate(
      cum_p_total = paste0(round(cumsum(total_p) * 100, 2), "%"),
      total_p     = paste0(round(total_p * 100, 2), "%")
    ) %>%
    dplyr::select(type, f, n, p, cum_p, total_p, cum_p_total) 
  
  # ── 4. Clean up factor level labels ───────────────────────────────────────────
  labels_full          <- table %>% dplyr::pull(f)
  max_break_appearance <- labels_full %>% as.character() %>% stringr::str_count("]") %>% max
  
  if (max_break_appearance == 1 && length(labels_full) > 0){
    splits     <- labels_full %>% stringr::str_split("] ", n = 2)
    labels     <- purrr::map_chr(splits, ~if (length(.x) >= 2) .x[2] else .x[1])
    value_part <- purrr::map_chr(splits, ~.x[1]) %>% stringr::str_remove("\\[")
    match      <- purrr::map2_lgl(labels, value_part, ~.x == .y)
    
    recoding        <- labels_full %>% as.character()
    new_names       <- dplyr::case_when(match ~ labels, !match ~ as.character(labels_full))
    names(recoding) <- new_names
    recoding        <- recoding[!duplicated(names(recoding))]
    
    table <- table %>% dplyr::mutate(f = forcats::fct_recode(f, !!!recoding))
  }
  
  names(table)[2] <- label
  
  # ── 5. Optional flextable formatting ─────────────────────────────────────────
  if (isTRUE(as_flextable)) {
    table <- .freq_format_table(table)
  }
  
  return(table)
}


#' Internal helper: apply flextable formatting to a freq_spss tibble
#'
#' @param tbl A tibble returned by the core freq_spss logic (before column rename).
#' @return A formatted flextable object.
#' @keywords internal

.freq_format_table <- function(tbl) {
  tbl %>%
    flextable::flextable() %>%
    flextable::autofit() %>%
    flextable::bg(i = ~type == "user_na", j = NULL, bg = "lightgrey") %>%
    flextable::bg(i = ~type != "user_na", j = NULL, bg = "lightblue") %>%
    flextable::bg(part = "header",                  bg = "lightgrey") %>%
    flextable::delete_columns("type") %>%
    flextable::set_header_labels(
      n           = "Frequency",
      p           = "Percentage",
      cum_p       = "Cumulative %",
      total_p     = "Total %",
      cum_p_total = "Total Cumulative %"
    ) %>%
    flextable::align(j = -1, align = "right", part = "all") %>%
    flextable::vline(j = c(1, 2, 4), part = "all")
}

```
<br> 

## Trials

### Basic application

```{r}

freq_spss(data = df, var = skipmeal)

```

<br> 

### With formatting

```{r}

freq_spss(data = df, var = skipmeal, as_flextable = TRUE)

```


```{r}
freq_spss(data = df, var = Poverty1, as_flextable = TRUE)
```

<br> 

### Illustration of what happens when some values have no labels...

```{r}

freq_spss(data = df, var = libauth, as_flextable = T)

```

<br> 

## Write function

```{r}

# Collect both functions as named list
fns <- list(
  freq_spss       = freq_spss,
  .freq_format_table = .freq_format_table
)

# Write each to the same file, separated by a blank line
writeLines(
  unlist(lapply(names(fns), function(nm) {
    c(paste0(nm, " <- "), deparse(fns[[nm]]), "")
  })),
  "/Users/joecrowley/R/Exploratory/Scripts/freq_spss.R"
)

```



