---
title: "Overview of SPSS data"
format: html
editor: visual
---

## Package set up and load data

```{r packages and data, collapse = T}

library(here)
library(haven)
library(tidyverse)
library(labelled)
library(flextable)

# Load BSA teaching data
df <- readRDS(here::here("data", "synthetic_bsa2019.rds"))
source(here::here("Scripts", "freq_spss.R"))

# Add missing value label for NatFrEst = 0
labelled::val_labels(df$NatFrEst) <- c(labelled::val_labels(df$NatFrEst), c("Zero people broken law" = 0))

vars <- names(df)
labels <- df %>% var_label(unlist = T, null_action = "fill")

```

<br>

## Contents

```{r contents, results='asis', echo = T}

for (xxx in 1:length(vars)){
  # print(xxx)
  cat(paste0(" * [",vars[xxx],"]"))
  cat(" :- ")
  cat(labels[xxx])
  cat("\n")

}

```

<br>

## Variable overview

Each variable is presented below. Variables with 50 or fewer unique values are shown as a frequency table. Variables with more than 50 unique values are tested for numeric conversion: if fewer than 5% of values fail to convert, they are shown as a histogram with any non-numeric values tabulated separately; otherwise they are shown as a frequency table. Where a variable has more than 10 unique values and all convert cleanly to numeric, a histogram is shown alongside the frequency table.

```{r variable overview, results='asis', eval = T, echo=T, message=F, warning=F}

for (var in vars) {

  # Get the variable label before any conversion, falling back to the variable name
  var_lbl         <- labelled::var_label(df[[var]]) %||% var

  # Convert to factor. nolabel_to_na = FALSE keeps values without labels as their
  # numeric codes rather than NAs, so numeric variables with a few labelled values
  # (e.g. a sentinel 0 = "Zero people broken law") don't lose most of their data.
  variable        <- df[[var]] %>% labelled::to_factor(user_na_to_na = TRUE, nolabel_to_na = FALSE)

  # Count unique non-missing values after factor conversion
  n_unique        <- variable %>% na.omit() %>% unique() %>% length()

  # Try converting factor levels to numeric. Labelled values produce strings like
  # "[0] Zero people broken law" which fail as.numeric() and become NA.
  # Unlabelled values produce plain numeric strings like "23" which convert fine.
  # pct_new_na is the share of originally non-missing values that fail conversion.
  numeric_attempt <- suppressWarnings(as.numeric(as.character(variable)))
  original_non_na <- !is.na(variable)
  new_na          <- is.na(numeric_attempt) & original_non_na
  pct_new_na      <- sum(new_na) / sum(original_non_na) * 100

  cat("\n")
  cat("[Contents] ")
  cat("\n\n\n")
  cat("### ", var, "\n")

  # Branch 1: ≤50 unique values (likely categorical) OR ≥5% fail numeric conversion.
  # Always show a freq_spss table. Also show a histogram if the variable has more
  # than 10 unique values and fewer than 5% fail numeric conversion —
  # i.e. it is essentially numeric but has some labelled values worth tabulating.
  if (n_unique <= 50 || pct_new_na >= 5) {

    if (n_unique > 10 && pct_new_na < 5) {
      p <- ggplot(data.frame(x = numeric_attempt), aes(x = x)) +
        geom_histogram(bins = 30) +
        labs(x = var, title = var_lbl) +
        theme_minimal()
      print(p)
    }

    freq_spss(data = df, var = !!rlang::sym(var), as_flextable = TRUE) %>%
      flextable_to_rmd()

  # Branch 2: >50 unique values AND <5% fail conversion — treat as continuous.
  # Show a histogram. If any values failed numeric conversion (non-numeric labels),
  # tabulate those separately below the histogram.
  } else {

    p <- ggplot(data.frame(x = numeric_attempt), aes(x = x)) +
      geom_histogram(bins = 30) +
      labs(x = var, title = var_lbl) +
      theme_minimal()
    print(p)

    freq_spss(data = df, var = !!rlang::sym(var), as_flextable = TRUE) %>%
      flextable_to_rmd()

  }

  cat("\n")
  cat("<br>")
  cat("\n")

}

```
