---
title: "Styled mschart charts"
format: html
editor: visual
---

## Overview

`deck` ships four chart builders that produce **native, editable Office
charts** (via `mschart`) in the NatCen house style, plus the house colour
palette. Each returns an `ms_*` object you feed straight into
`full_width_chart_slide()` or `text_and_chart_slide()`.

They were ported from the old `Visualisations/old/mscharts code, functions.R`
scratch script -- the colour-selection block that was copy-pasted into four
functions, and the house-style theme repeated in every one, are now the
single shared helpers `deck:::.chart_colours()` and
`deck:::.natcen_chart_style()`.

```r
devtools::install("deck")   # or devtools::load_all("deck")
```

<br>

```{r packages, collapse = T}
library(tidyverse)
library(mschart)
library(officer)
library(deck)

set.seed(1)
```

<br>

## The house palette

`nc_palette()` returns a 5-hue by 5-shade matrix of hex colours.

```{r}
nc_palette()
```

The `colour` argument of every builder accepts:

-   a **hue name** -- `"green"`, `"purple"`, `"pink"`, `"blue"`, `"orange"` -- for successive shades of that hue
-   `"categorical"` (5 colours) or `"categorical10"` (10) -- distinct hues
-   a single colour (hex or R colour name)
-   a vector of colours

<br>

## Test data

```{r}
bars <- data.frame(
  category = c("A", "B", "C", "D"),
  value    = c(10, 23, 15, 8)
)

grouped <- data.frame(
  meta_cat = factor(rep(c("One", "Two", "Three"), 4), levels = c("One", "Two", "Three")),
  category = rep(c("A", "B", "C", "D"), each = 3),
  value    = c(10, 9, 3, 23, 10, 6, 15, 11, 25, 8, 17, 23)
)

monthly <- data.frame(
  month  = rep(seq(as.Date("2020-01-01"), by = "month", length.out = 12), 4),
  series = rep(c("A", "B", "C", "D"), each = 12),
  value  = round(runif(48, 20, 80))
)
```

```{r}
ppt <- mint_pptx()
```

<br>

## Bar chart

Flat fill (default NatCen green):

```{r}
ppt <- ppt |>
  full_width_chart_slide(
    title = "Bar chart",
    chart = bar_chart(bars, x = "category", y = "value")
  )
```

One colour per bar, horizontal:

```{r}
ppt <- ppt |>
  full_width_chart_slide(
    title = "Bar chart - one colour per bar, horizontal",
    chart = bar_chart(bars, x = "category", y = "value",
                      colour = "categorical", horizontal = TRUE)
  )
```

<br>

## Clustered bar chart

```{r}
ppt <- ppt |>
  full_width_chart_slide(
    title = "Clustered bar chart",
    chart = clustered_bar_chart(grouped, x = "category", y = "value",
                                group = "meta_cat")
  )
```

<br>

## Stacked bar chart

`percent = TRUE` caps the value axis at 100; `num_fmt` sets the label / axis
number format.

```{r}
ppt <- ppt |>
  full_width_chart_slide(
    title = "Stacked bar chart",
    chart = stacked_bar_chart(grouped, x = "category", y = "value",
                              group = "meta_cat", horizontal = TRUE,
                              colour = "categorical10")
  )
```

<br>

## Line chart

```{r}
ppt <- ppt |>
  full_width_chart_slide(
    title = "Line chart",
    chart = line_chart(monthly, x = "month", y = "value", group = "series")
  )
```

<br>

## Save

```{r}
print(ppt, target = "Example styled mscharts.pptx")
```
