---
title: "explo: function reference and usage"
format: html
editor: visual
---

## What this is

This document demonstrates the functions in the `explo` package (formerly `QuickDescriptives`). Unlike the old workflow (functions defined *inside* Quarto chunks, then `writeLines(deparse(...))`'d out to loose `.R` files), the functions here live in `../R/*.R` as ordinary package code with roxygen documentation. This `.qmd` only *uses* the package -- it doesn't define anything.

While the package is still a pilot (not yet installed), `devtools::load_all()` loads it straight from source, same as `library()` would once installed.

```{r setup}
#| collapse: true
library(tidyverse)
library(haven)
library(labelled)
devtools::load_all("..")

# Reuse the same synthetic dataset the Exploratory book uses.
df <- readRDS(file.path("..", "..", "data", "synthetic_bsa2019.rds"))
```

<br>

## Multiple frequencies

### `count_across()`

Stacks frequency counts for several variables into one long table.

```{r}
df |> count_across(vars = c("skipmeal", "Politics", "Poverty1"))
```

## Crosstabs

### `ctab()`

Two-way crosstab of one variable against a breakdown variable.

```{r}
ctab(df, skipmeal, RSex)
```

### `ctabs()` + `ctabs_xlsx()`

Crosstabs several outcomes against several predictors at once, exportable to a formatted Excel workbook. `ctabs()` always returns an `overview` element too -- marginal frequencies (via `count_across()`) for every variable involved, outcomes and predictors alike -- alongside the `tables` element holding the actual crosstabs. This isn't optional: `ctabs_xlsx()` always writes it as the first ("Overview") sheet in the workbook.

```{r}
tbl <- df |> ctabs(vars = c("HEdQual3"), preds = c("RClassGp", "RAgeCat"))
tbl$overview
tbl$tables[[1]][[1]]
```

```{r}
#| eval: false
ctabs_xlsx(tbl, filename = "ctabs_export.xlsx")
```

### `mfreq()`

Frequency table across several variables (works best when they share a common response scale; shown here with two unrelated categoricals just to demonstrate the mechanics).

```{r}
df |> select(Married, ChildHh) |> mfreq(labels = TRUE)
```

<br>

## Missing data

### `df_miss()`

```{r}
df |> df_miss() |> arrange(desc(p_miss)) |> head()
```

### `df_miss_brk()`

Missingness broken down by a grouping variable, formatted as a flextable.

```{r}
df |> df_miss_brk(brk = RAgeCat, type = "p_miss", form = "flx")
```

<br>

## Numeric summaries

### `num_sum()`

```{r}
df |> num_sum(vars = c("libauth", "leftrigh", "welfare2"))
```

### `num_crs()`

```{r}
num_crs(df, nums = c("libauth", "leftrigh", "welfare2"), brk = "RAgeCat", stats = "mean")
```

<br>

## Predictor-level breakdowns

### `natural_breaks()`

```{r}
natural_breaks(df, vars = c("libauth", "leftrigh", "welfare2"), nbrk = 5)
```

### `return_one2()` + `return_one2_format()`

```{r}
return_one2(df, out = "skipmeal", lev = c("n_val", "p_val"), preds = c("RSex", "RAgeCat")) |>
  return_one2_format(print = "f")
```
